Bird beak's elaboration and exploration

### 2021-10-11 Analysis/03.1-results.Rmd
Update 03.1-results.Rmd

```{r, fig.height = 20, fig.width = 10, echo = FALSE, eval = FALSE}
## Run the analyses between groups
super_results_between <- covar.projections.wrapper(superorder_shapespace,
                                             type = "between",
                                             base = "phylogeny",
                                             dispRity.out = TRUE)
## Run the analyses between groups
order_results_between <- covar.projections.wrapper(order_shapespace,
                                             type = "between",
                                             base = "phylogeny",
                                             dispRity.out = TRUE)
## Run the analyses of groups on the phylogeny
super_results_phylo <- covar.projections.wrapper(superorder_shapespace,
                                            type = "within",
                                            base = "phylogeny",
                                            incl.base = TRUE,
                                            dispRity.out = TRUE,
                                            verbose = TRUE)
## Run the analyses of groups on the phylogeny
order_results_phylo <- covar.projections.wrapper(order_shapespace,
                                            type = "within",
                                            base = "phylogeny",
                                            incl.base = TRUE,
                                            dispRity.out = TRUE,
                                            verbose = TRUE)
## Run the analyses groups on themselves
super_results_within <- covar.projections.wrapper(superorder_shapespace,
                                            type = "within",
                                            dispRity.out = TRUE,
                                            verbose = TRUE)
## Run the analyses groups on themselves
order_results_within <- covar.projections.wrapper(order_shapespace,
                                            type = "within",
                                            dispRity.out = TRUE,
                                            verbose = TRUE)
## Save all
allbirds_8748t_8Dshapespace_31levels_results_list <- list("super_results_between" = super_results_between,
                     "order_results_between" = order_results_between,
                     "super_results_phylo"   = super_results_phylo,
                     "order_results_phylo"   = order_results_phylo,
                     "super_results_within"  = super_results_within,
                     "order_results_within"  = order_results_within)
save(allbirds_8748t_8Dshapespace_31levels_results_list, file = "../Data/Processed/allbirds_8748t_8Dshapespace_31levels_results_list.rda")
```

@@

    ## Which will be the centre plot?
    centre <- 1
    ## Plot everything
    op <- par(mfrow = c(n, 1), mar = c(7, 4, 4, 2) + 0.1)
    for(i in 1:n) {
        plot(results[[i]], col = col, main = ifelse(i == centre, main, ""), las = 2, xlab = "")
    }
    par(op)
}
wrap.plot(super_results_between,
          col = rainbow(n.subsets(superorder_shapespace)-1),
          main = "Superorders differences")
```
Figure: these results show the projections ("position"), rejection ("distance") and angles ("degree") of the posteriors of each group's major axes onto the corresponding global phylogenetic major axes. For each panel, the highest the score the more the clade is elaborating, exploring, exploring perpendicular to the main axes for the position, distance and degree respectively.

```{r, fig.height = 20, fig.width = 10, echo = FALSE}
wrap.plot(order_results_between,
          col = rainbow(n.subsets(order_shapespace)-1),
          main = "Superorders differences")
```

@@

## Plot the results
wrap.plot(super_results_phylo,
          col = rainbow(n.subsets(superorder_shapespace)-1),
          main = "Projections per superorder\non the global phylogeny axes")
```
Figure: the projection of each species per group onto the posterior distribution of global major phylogenetic axis.

@@

## Plot the results
wrap.plot(order_results_phylo,
          col = rainbow(n.subsets(order_shapespace)-1),
          main = "Projections per order\non the global phylogeny axes")
```

@@

## Plot the results
wrap.plot(super_results_within,
          col = rainbow(n.subsets(superorder_shapespace)-1),
          main = "Projections per superorder\nonto each superorder")
```
Figure: the projection of each species per group onto the posterior distribution of each clade's major phylogenetic axis (the grey distributions corresponds to the projections of all the species, irregardless of their group, onto the global phylogenetic major axes).

@@

## Plot the results
wrap.plot(order_results_within,
          col = rainbow(n.subsets(order_shapespace)-1),
          main = "Projections per order\nonto each order")
```

>>>>

```{r, fig.height = 20, fig.width = 10, echo = FALSE, eval = TRUE}
## Run the analyses between groups
super_results_between <- covar.projections.wrapper(superorder_shapespace,
                                             type = "between",
                                             base = "phylogeny",
                                             dispRity.out = TRUE)
## Run the analyses between groups
order_results_between <- covar.projections.wrapper(order_shapespace,
                                             type = "between",
                                             base = "phylogeny",
                                             dispRity.out = TRUE)
## Run the analyses of groups on the phylogeny
super_results_phylo <- covar.projections.wrapper(superorder_shapespace,
                                            type = "within",
                                            base = "phylogeny",
                                            incl.base = TRUE,
                                            dispRity.out = TRUE,
                                            verbose = TRUE)
## Run the analyses of groups on the phylogeny
order_results_phylo <- covar.projections.wrapper(order_shapespace,
                                            type = "within",
                                            base = "phylogeny",
                                            incl.base = TRUE,
                                            dispRity.out = TRUE,
                                            verbose = TRUE)
## Run the analyses groups on themselves
super_results_within <- covar.projections.wrapper(superorder_shapespace,
                                            type = "within",
                                            dispRity.out = TRUE,
                                            verbose = TRUE)
## Run the analyses groups on themselves
order_results_within <- covar.projections.wrapper(order_shapespace,
                                            type = "within",
                                            dispRity.out = TRUE,
                                            verbose = TRUE)
## Save all
allbirds_8748t_8Dshapespace_31levels_results_list <- list(
        "super_results_between" = super_results_between,
        "order_results_between" = order_results_between,
        "super_results_phylo"   = super_results_phylo,
        "order_results_phylo"   = order_results_phylo,
        "super_results_within"  = super_results_within,
        "order_results_within"  = order_results_within)
save(allbirds_8748t_8Dshapespace_31levels_results_list, file = "../Data/Processed/allbirds_8748t_8Dshapespace_31levels_results_list.rda")
```

@@

    ## Which will be the centre plot?
    centre <- 1
    ## Plot everything
    op <- par(mfrow = c(n, 1), mar = c(7, 4, 4, 2) + 0.1)
    for(i in 1:n) {
        plot(results[[i]], col = col, main = ifelse(i == centre, main, ""), las = 2, xlab = "")
    }
    par(op)
}
wrap.plot(results$super_results_between,
          col = rainbow(n.subsets(superorder_shapespace)-1),
          main = "Superorders differences")
```
Figure: these results show the projections ("position"), rejection ("distance") and angles ("degree") of the posteriors of each group's major axes onto the corresponding global phylogenetic major axes. For each panel, the highest the score the more the clade is elaborating, exploring, exploring perpendicular to the main axes for the position, distance and degree respectively.

```{r, fig.height = 20, fig.width = 10, echo = FALSE}
wrap.plot(results$order_results_between,
          col = rainbow(n.subsets(order_shapespace)-1),
          main = "Superorders differences")
```

@@

```{r, fig.height = 20, fig.width = 10, echo = FALSE, eval = FALSE}
## Plot the results
wrap.plot(results$super_results_phylo,
          col = rainbow(n.subsets(superorder_shapespace)-1),
          main = "Projections per superorder\non the global phylogeny axes")
```
Figure: the projection of each species per group onto the posterior distribution of global major phylogenetic axis.

@@

## Plot the results
wrap.plot(results$order_results_phylo,
          col = rainbow(n.subsets(order_shapespace)-1),
          main = "Projections per order\non the global phylogeny axes")
```

@@

## Plot the results
wrap.plot(results$super_results_within,
          col = rainbow(n.subsets(superorder_shapespace)-1),
          main = "Projections per superorder\nonto each superorder")
```
Figure: the projection of each species per group onto the posterior distribution of each clade's major phylogenetic axis (the grey distributions corresponds to the projections of all the species, irregardless of their group, onto the global phylogenetic major axes).

@@

## Plot the results
wrap.plot(results$order_results_within,
          col = rainbow(n.subsets(order_shapespace)-1),
          main = "Projections per order\nonto each order")
```

## Different scores distributions per species

```{r, eval = TRUE, echo = FALSE, eval = FALSE}
library(hrdcde)

#@param results the results list
#@param type which type (position, distance or angle)
#@param cent.tend which central tendency fun (default is median)
#@param use.hdr whether to use kernel density estimate for the outputs (default is TRUE)
#@param prob which quantiles to use (default is c(50, 95, 99))
#@param ... plot options

get.distributions <- function(results, type, cent.tend = median, use.hdr = TRUE, prob = c(50, 95, 99), ...) {

    ## Extracting all the results
    disparity <- lapply(results[[type]]$disparity, function(X, type) return(X$elements))

    ## Recentre the results
    if(type == "position") {
        disparity <- lapply(disparity, function(X) return(X-0.5))
    }

    ## Measure the central tendencies
    centrals <- lapply(disparity, function(X, fun) apply(X, 1, fun), fun = cent.tend)
    ## Number of groups
    n_groups <- length(centrals)

    ## Calculate and plot the hdrcde
    op_tmp <- par(mfrow = c(ceiling(sqrt(n_groups)), round(sqrt(n_groups))))
    densities <- list()
    for(i in 1:length(centrals)) {
        if(use.hdr) {
            densities[[i]] <- hdr.den(centrals[[i]], prob = prob, main = names(centrals)[i], ...)
        } else {
            hist(centrals[[i]], main = names(centrals)[i], ...)
        }
    }
    par(op_tmp)

    ## Find the extreme species
    get.extremes.hdr <- function(density, central) {
        ## Get the minimums and maximums
        mins <- which(central <= density$hdr[1,1])
        maxs <- which(central >= density$hdr[1,2])
        ## Get their values
        return(c(central[names(mins)], central[names(maxs)]))
    }
    CI.converter <- function(CI) {
        sort(c(50-CI/2, 50+CI/2)/100)
    }
    get.extremes.quantile <- function(central, prob) {
        ## Get the quantiles
        quants <- quantile(central, probs = CI.converter(max(prob)))
        ## Get the minimums and maximums
        mins <- which(central <= quants[1])
        maxs <- which(central >= quants[2])
        ## Get their values
        return(c(central[names(mins)], central[names(maxs)]))
    }

    if(use.hdr) {
        output <- mapply(get.extremes.hdr, densities, centrals, SIMPLIFY = FALSE)
        names(output) <- names(centrals)
    } else {
        output <- lapply(centrals, get.extremes.quantile, prob = prob)
    }

    ## Return the as a table
    for(i in 1:length(output)) {
        output[[i]] <- data.frame("level" = rep(names(output)[i], length(output[[i]])),
                                  "species" = names(output[[i]]),
                                  "score"   = unname(output[[i]]))
    }
    return(output)
}
```

```{r, eval = TRUE, echo = FALSE, fig.height = 12, fig.width = 12, eval = FALSE}
get.distributions(results$super_results_between, type = "position")
```
